Fair, Stable Exchange Networks through Quenched Merchant Location and Idiosyncratic Trading Costs

### 2019-06-25 writing/Paper/main.tex
minor changes in document

In this article, we show that such monopolies can be prevented by introducing an idiosyncratic buyer-seller transaction cost.
This tax leverages the benefits of modern distribution channels, making sure that no buyer remains disadvantaged. 
At the same time, innovative products can spread globally, whereas less specialized products remain local. 
This is achieved through the following taxation mechanism: 
Consider a set of $|F|$ firms, all offering the same product, and a set of $|B|$ potential buyers (customers). 
For any buyer $b \in B$ and firm $f \in F$, we define the ordinal distance $D = D(b,f)$ such that the firm $f$ that is physically closest to $b$ has distance $D(b,f)=1$, the next closest has distance $D=2$, and the $n$-th closest has distance $D=n$. 
The tax when customer $b$ buys from the $n$-th closest firm is then proportional to $D^\gamma$, for some exponent $\gamma > 0$. 
If we were to set $\gamma = 0$, everybody can buy from every firm at single tax rate. 
If we were to set $\gamma=\infty$, a seller can only buy from its closest firm.
A gamma in-between those two extremes represents a trade-off: 
Close-by products are taxed less than far away products. 
However, because the distance is ordinal, absolute distances do not matter. 
No matter whether the next closest firm is $100$ meters or $100$ kilometres, the taxation is the same. 
In particular, if there is only one firm offering the product, everybody buys for the same tax (since $D=1$ for every buyer), enabling innovations to spread globally.

Through tuning of $\gamma$ and targeted redistribution of tax money to regions of low economic activity, we show that this model is able to fuel a self-consistent economy, 
with fair competition, where innovations can spread freely, but without the chance of monopolies emerging through disproportional feedback mechanisms. 
In summary, we present a model that avoids the winner-take-all monopoly tendency and associated negative externalities of today's Internet enabled business through a tax based on a topological constraint \green{the tax is not always small, and definitely not in our examples}.

>>>>

In this article, we show that such monopolies can be prevented by introducing an idiosyncratic buyer-seller transaction cost.
This tax leverages the benefits of modern distribution channels, making sure that no buyer remains disadvantaged. 
At the same time, innovative products can spread globally, whereas less specialized products remain local. 
This is achieved through the following taxation mechanism: 
Consider a set of $|F|$ firms, all offering the same product, and a set of $|B|$ potential buyers (customers). 
For any buyer $b \in B$ and firm $f \in F$, we define the ordinal distance $D = D(b,f)$ such that the firm $f$ that is physically closest to $b$ has distance $D(b,f)=1$, the next closest has distance $D=2$, and the $n$-th closest has distance $D=n$. 
The tax when customer $b$ buys from the $n$-th closest firm is then proportional to $D^\gamma$, for some exponent $\gamma > 0$. 
If we were to set $\gamma = 0$, everybody can buy from every firm at the same tax rate. 
If we were to set $\gamma=\infty$, a seller can only buy from its closest firm.
A gamma in-between those two extremes represents a trade-off: 
Close-by products are taxed less than far away products. 
However, because the distance is ordinal, absolute distances do not matter. 
No matter whether the next closest firm is $100$ meters or $100$ kilometres, the taxation is the same. 
In particular, if there is only one firm offering the product, everybody buys for the same tax (since $D=1$ for every buyer), enabling innovations to spread globally.

Through tuning of $\gamma$ and targeted redistribution of tax money to regions of low economic activity, we show that this model is able to fuel a self-consistent economy, 
with fair competition, where innovations can spread freely, but without the chance of monopolies emerging through disproportional feedback mechanisms. 
In summary, we present a model that avoids the winner-take-all monopoly tendency and associated negative externalities of today's Internet enabled business through a discretised, distance-based tax.
